---
title: "ESM 206 Lab 5"
author: "Kazia Mermel"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to parse dates with 'lubridate' package
- Get counts of observations with count()
- Uncount() rows
- One and two-sample t-test (we're ahead in lab of where we are in lecture)
- Create a heatmap with geom_tile()

```{r, include = FALSE}
# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>%
  janitor::clean_names()
  
```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)
```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}
ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram per site
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

# quantile-quantile

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)

```

#### Convert 'date' column to class 'Date'

We'll use the lubridate package to convert to date format, and then to help us easily parse month and year

```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Now let's parse year and month using lubridate::month() and lubridate::year()

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Now let's find counts of observed lobsters based on different hierarchical groupings.

First, let's count lobsters by year and month.

```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

lobster_site
```

If we want to create a summary table that contains statistics other than counts by group, it's easier to use group_by() plus the n() function

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

lobster_summary
```

#### Find confidence intervals

Use t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples (....this will be covered conceptually in lectures week 6)

```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)

ivee_lobsters

```

#### Two-sample t-test to compare means

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a two-sample t-test for means comparison is appropriate.

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

mn_ttest
```

There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p < 0.001, alpha = 0.05). 
(This should always be the least interesting part of a description of data. More interesting? What are the actual means, what's the structure of the data, etc.)

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~site, data = lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic

```

Here is the p-value: `r mn_ttest2$p.value`

Here is my t-statistic: `r mn_ttest2$statistic`

Both of those sentences are bad ways of talking about stats but are serving to show how to format so that the stat values will auto-update

#### Now, a heat map!

```{r}
lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)


ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

